---
title: "Live Session Unit 4 Assignment"
author: "Alec Neff"
date: "6/9/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. FiveThirtyEight Data

a. Install the fivethirtyeight package.
```{r,echo=TRUE}
install.packages("fivethirtyeight",repos = "https://cran.rstudio.com/")
```

b. In the listing of Data sets in package 'fivethirtyeight', assign the eighteenth data set to an object 'df'.
```{r,echo=TRUE}
data(package="fivethirtyeight")
data("college_recent_grads",package = "fivethirtyeight")
df<-college_recent_grads
```

c. Use a more detailed list of the data sets to write out the URL in a comment to the related news story.

The data can be found at: https://github.com/rudeboybert/fivethirtyeight/blob/master/data/college_recent_grads.rda

d. Using R command(s), give the dimensions and column names of this data frame.
```{r,echo=TRUE}
str(df)
```


## 2. Data Summary

a. Write an R command that gives you the column names of the data frame.  Right after that, write one that counts the number of columns but not rows. Hint: the number should match one of your numbers in Question 1d for dimensions.
```{r,echo=TRUE}
colnames(df)
ncol(df)
```

b. Generate a count of each unique major_category in the data frame.  I recommend using libraries to help.  I have demonstrated one briefly in live-session.  To be clear, this should look like a matrix or data frame containing the major_category and the frequency it occurs in the dataset. Assign it to major_count.
```{r,echo=TRUE}
major_count<-table(df$major_category)
major_count
```
or
```{r,echo=TRUE}
as.matrix(major_count)
```
c. To make things easier to read, enter par(las=2) before your plot to make the text perpendicular to the axis.  Make a barplot of major_count.  Make sure to label the title with something informative (check the vignette if you need), label the x and y axes, and make it any color other than grey.  Assign the major_category labels to their respective bars.  Flip the barplot horizontally so that bars extend to the right, not upward.  All of these options can be done in a single pass of barplot().  Note: it's okay if it's wider than the preview pane.
*I think labelling the y axis here is problematic to appearances and is superfluous as the bar names serve are self-explanatory considering Major Category is included in the title.
```{r,echo=TRUE}
par(las=1,mar=c(5.1,15.1,4.1,3.1))
barplot(as.matrix(major_count)[16:1],main = "Number of Polled Graduates by Major Category",xlab = "Count",horiz = TRUE,col = "darkgreen",names.arg = row.names(major_count)[16:1],xlim = c(0,30))
```

d. Write the fivethirtyeight data to a csv file. Make sure that it does not have row labels.
```{r,echo=TRUE}
write.csv(df,file = "/Users/acneff/Documents/Data_Science/School/DDS/HW4/fte_college_recent_grads.csv",row.names = FALSE)
```

## 3. Codebook
a-c: see link in d

d. In your RMarkdown script, please provide the link to this GitHub so the grader can see it.

https://github.com/acneff/smu_msds_hw

## Thanks for joining me. Come back when you can stay longer!