---
title: "Live Session Unit 5 Assignment"
author: "Alec Neff"
date: "6/14/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
*Reader's Note: While the text of the questions warns that the data in the text files is not comma-separated values (.csv), I found it elementary to add column names to the files, do some find/replace magic on the semicolons in the yob2016.txt file, and save both as .csv files.*

# Question 1 - Data Munging
Utilize **yob2016.txt** for this question.  This file is a series of popular children’s names born in the year 2016 in the United States.  It consists of three columns with a first name, a gender, and the amount of children given that name.  However, the data is raw and will need cleaning to make it tidy and usable.

### 1A
First, import the .txt file into R so you can process it.  Keep in mind this is not a CSV file.  You might have to open the file to see what you’re dealing with.  Assign the resulting data frame to an object, **df**, that consists of three columns with human-readable column names for each.

```{r,echo=TRUE}
df<-read.csv(file = "/Users/acneff/Documents/Data_Science/School/DDS/HW5/yob2016.csv")
```

### 1B
Display the summary and structure of **df**.

```{r,echo=TRUE}
summary(df)
str(df)
```

### 1C
Your client tells you that there is a problem with the raw file.  One name was entered twice and misspelled.  The client cannot remember which name it is; there are thousands he saw! But he did mention he accidentally put three y’s at the end of the name.  Write an R command to figure out which name it is and display it.

```{r,echo=TRUE}
grep("yyy",df$name,value = TRUE)
```

### 1D
Upon finding the misspelled name, please remove this particluar observation, as the client says it's redundant.  Save the remaining dataset as an object: **y2016**.
```{r,echo=TRUE}
grep("yyy",df$name)
df[212,]
y2016<-df[-212,]
```

# 2 Data Merging
Utilize **yob2015.txt** for this question.  This file is similar to yob2016, but contains names, gender, and total children given that name for the year 2015.

### 2A
Like 1A, please import the .txt file into R.  Look at the file before you do.  You might have to change some options to import it properly.  Again, please give the dataframe human-readable column names.  Assign the dataframe to **y2015**.
```{r,echo=TRUE}
y2015<-read.csv(file = "/Users/acneff/Documents/Data_Science/School/DDS/HW5/yob2015.csv")
```

### 2B
Display the last ten rows in the dataframe.  Describe something you find interesting about these 10 rows.
```{r,echo=TRUE}
tail(y2015,n=10)
```
*Every one of these arguably strange names was given to five boys each.  Every.  One.  Also, just some predictive advice, don't mess with the Zohar.*

### 2C
Merge **y2016** and **y2015** by your Name column; assign it to **final**.  The client only cares about names that have data for both 2016 and 2015; there should be no NA values in either of your amount of children rows after merging.
```{r,echo=TRUE}
final<-merge(y2016,y2015,by="name")
```
*Let's make sure there are no NAs.*
```{r,echo=TRUE}
which(is.na(final$count.x)==TRUE)
which(is.na(final$count.y)==TRUE)
```

# 3 Data Summary
Utilize your data frame object **final** for this part.

### 3A
Create a new column called "Total" in **final** that ads the *amount of children* in 2015 and 2016 together.  In those two years combined, how many people were given popular names?
```{r,echo=TRUE}
final$total<-final$count.x + final$count.y
sum(final$total)
```
Over 11.4 million.  Comparable to the current population of Guangzhou, China.

### 3B
Sort the data by Total.  What are the top 10 most popular names?
```{r,echo=TRUE}
head(final[order(-final$total),],n=10)
```

### 3C
The client is expecting a girl! Omit boys and give the top 10 most popular girl's names.
*I wonder...some names are unisex and you never know if one is listed for boys in one year and girls the next...it can happen.*
```{r,echo=TRUE,R.options=list(max.print=50)}
which(final$gender.x != final$gender.y)
```
*There are quite a few.  It's probably easier to get a new dataframe with only the entries that post "F" in the sex column for either year, and find the top ten from that.*
```{r,echo=TRUE}
final_girls<-subset(final,final$gender.x=="F"|final$gender.y=="F")
head(final_girls[order(-final_girls$total),],n=10)
```
*It turns out that worrying about gender-switching names was unnecessary in our investigation, but the action required to be certain was simple.*

### 3D
Write these top 10 girl names and their Totals to a CSV file.  Leave out the other columns entirely.
*I'm going to start by creating a object with those top ten names to simplify the write.csv() statement.*
```{r,echo=TRUE}
topten_girls<-head(final_girls[order(-final_girls$total),],n=10)
write.csv(topten_girls[,c("name","total")],file = "/Users/acneff/Documents/Data_Science/School/DDS/HW5/topten_girls.csv",row.names = FALSE)
```

# 4 Upload to GitHub
Push at minimum your RMarkdown for this homework assignment and a Codebook to one of your GitHub repositories (you might place this in a Homework repo like last week).  The Codebook should contain a short definition of each object you create, and if creating multiple files, which file it is contained in.  You are welcome and encouraged to add other files—just make sure you have a description and directions that are helpful for the grader.

*Link to GitHub repo folder for this assignment:*
[Live Session Unit 5 Assignment](https://github.com/acneff/smu_msds_hw/tree/master/6306/HW5)

## Thanks for stopping by; come back when you can stay longer!